Supporting Dynamic Data Partitioning in Distributed Memory Architectures with Logical Regions

### 2012-11-04 doc/types-pldi2013/intro.tex
intro

To support additional parallelism in the presence of aliased data,
Legion provides {\em coherence} modes at the granularity of regions, which allow
relaxed constraints on the execution order of tasks, even in case where they
do access the same region.  The programmer may
specify that tasks conflicting on a region  must execute serially (in program
order, the default), {\em atomically} with respect to each other, or even
{\em simultaneously} if the application has alternative methods for ensuring
determinism or can tolerate non-deterministic behavior in that region..

To guarantee the safety and correctness of programs, Legion relies on
a combination of static and dynamic analysis based on the privilege
and coherence systems.  
%To efficiently guarantee the safety and
%correctness of programs using first-class regions,
%privilege and coherence properties in the presence of arbitrary region
%aliasing, Legion relies on a combination of static and
%dynamic checks.  
The critical feature in Legion that makes static analysis tractable
and dynamic analysis efficient in the presence of potential aliasing is logical regions.
Alias analysis is a very hard static analysis problem, but is both easy and
inexpensive if done dynamically at the granulatiry of logical regions instead
of individual memory locations; similarly
checks on individual memory accesses are expensive if done dynamically, but
cheap if done statically within the context of local logical region privileges.
%A key theme that runs through the design is that 
%alias analysis is a very hard static analysis problem, but is both easy and
%inexpensive if done dynamically at the granularity of regions instead of individual
%heap locations, while checks that must be done on individual values are only
%cheap if done statically.  
Explicating and evaluating the design of Legion's
static and dynamic semantics and its implications are the topics of this paper,
which we present as a series of contributions:
\begin{itemize}
\item We present a type system for {\em Core Legion} programs that statically
  verifies the safety of individual pointers and region
  privileges at task call boundaries (Section~\ref{subsec:coretypes}).
%  A subtle issue arises in checking privileges when regions are
%  stored in and later retrieved from heap data structures.  The type
%  system provides enough static information that the Legion runtime
%  can leverage its dynamic information about region aliasing to
%  perform dynamic privilege checks guaranteeing safety.

>>>>

To support additional parallelism in the presence of aliased data,
Legion provides {\em coherence} modes at the granularity of regions,
which allow relaxed constraints on task execution order, even
in case where tasks do access the same region.  The programmer may
specify that tasks conflicting on a region must execute serially (in
program order, the default), {\em atomically} with respect to each
other, or even {\em simultaneously} if the application has alternative
methods for ensuring determinism or can tolerate non-deterministic
behavior in that region.

%To guarantee the safety and correctness of programs, Legion relies on
%a combination of static and dynamic analysis based on the privilege
%and coherence systems.  
%To efficiently guarantee the safety and
%correctness of programs using first-class regions,
%privilege and coherence properties in the presence of arbitrary region
%aliasing, Legion relies on a combination of static and
%dynamic checks.  
The critical feature in Legion that makes static analysis tractable
and dynamic analysis efficient in the presence of potential aliasing is logical regions.
Alias analysis is a very hard static analysis problem, but is both easy and
inexpensive if done dynamically at the granulatiry of logical regions instead
of individual memory locations; similarly
checks on individual memory accesses are expensive if done dynamically, but
cheap if done statically within the context of local logical region privileges.
%A key theme that runs through the design is that 
%alias analysis is a very hard static analysis problem, but is both easy and
%inexpensive if done dynamically at the granularity of regions instead of individual
%heap locations, while checks that must be done on individual values are only
%cheap if done statically.  
Explicating and evaluating the design of Legion's
static and dynamic semantics and its implications are the topics of this paper,
which we present as a series of contributions:
\begin{itemize}
\item We present a type system for {\em Core Legion} programs that statically
  verifies the safety of individual pointers and region
  privileges at task call boundaries (Section~\ref{subsec:coretypes}).
%  A subtle issue arises in checking privileges when regions are
%  stored in and later retrieved from heap data structures.  The type
%  system provides enough static information that the Legion runtime
%  can leverage its dynamic information about region aliasing to
%  perform dynamic privilege checks guaranteeing safety.
